Notes for ISIPTA poster

### 2016-07-12 asce-paper.tex
more in sec 2 (drop k there?)

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $n_k$ exchangeable components in the system.
For each type $k$ component, we assume for its lifetime $T_i^k$ ($i=1,\ldots,n_k$, $k = 1, \ldots, K$)
a Weibull distribution with fixed shape parameter $\beta_k > 0$, with density and cdf%
\footnote{Our approach would be possible also for other parametric lifetime distributions
that form a canonical exponential family
(see, e.g., \citeNP[p.~202 and 272f]{2000:bernardosmith}, or \citeNP[p.~8]{diss}).}
\begin{linenomath*}
\begin{align}
f_k(t_i^k \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_i^k)^{\beta_k-1} e^{-\frac{(t_i^k)^{\beta_k-1}}{\lambda_k}}\, \\
F_k(t_i^k \mid \lambda_k) &= 1 - e^{-\frac{(t_i^k)^\beta_k}{\lambda_k}} = P_k(T_i^k \leq t_i^k \mid \lambda_k)\,,
\end{align}
\end{linenomath*}
%For each $k$, component lifetimes ($i=1,\ldots,n_k$) are
in short $T_i^k \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$.
%where $\beta_k$ is fixed and known:

Conjugate prior is $\lambda_k \sim \ig(a_k,b_k)$:
\begin{linenomath*}
\begin{align}
f_{\lambda_k}(\lambda_k\mid a_k,b_k) &= \frac{(b_k)^{a_k}}{\Gamma(a_k)} \lambda_k^{-a_k -1} e^{-\frac{b_k}{\lambda_k}}
\end{align}
\end{linenomath*}
In terms of canonical parameters $\nkz$ and $\ykz$, we assume $\lambda_k \mid \nkz,\ykz \sim \ig(\nkz + 1, \nkz\ykz)$.

>>>>

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $n_k$ exchangeable components in the system.
For each type $k$ component, we assume for its lifetime $T_i^k$ ($i=1,\ldots,n_k$, $k = 1, \ldots, K$)
a Weibull distribution with fixed shape parameter $\beta_k > 0$,
in short $T_i^k \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$,
with density and cdf%
\footnote{Our approach would be possible also for other parametric lifetime distributions
that form a canonical exponential family
(see, e.g., \citeNP[p.~202 and 272f]{2000:bernardosmith}, or \citeNP[p.~8]{diss}).}
\begin{linenomath*}
\begin{align}
f_k(t_i^k \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_i^k)^{\beta_k-1} e^{-\frac{(t_i^k)^{\beta_k-1}}{\lambda_k}}\,, \\
F_k(t_i^k \mid \lambda_k) &= 1 - e^{-\frac{(t_i^k)^\beta_k}{\lambda_k}} = P_k(T_i^k \leq t_i^k \mid \lambda_k)\,,
\end{align}
\end{linenomath*}
where $\lambda_k > 0$ and $t > 0$.%
\footnote{The shape parameter $\beta_k$ determines whether the hazard rate is increasing ($\beta_k > 1$)
or decreasing ($\beta_k < 1$) over time.
For $\beta_k=1$, we obtain the Exponential distribution with constant hazard rate as a special case.
The value for $\beta_k$ will thus often be clear from practical considerations.}
The scale parameter $\lambda_k$ can be interpreted through the relation
\begin{linenomath*}
\begin{align}
\E[T_i^k \mid \lambda_k] &= \lambda_k^{1/\beta_k}\, \Gamma(1 + 1/\beta_k)\,.
\label{eq:lambdainterpret}
\end{align}
\end{linenomath*}
For encoding expert knowledge about the reliability of the components,
we need to assign a prior distribution over the scale parameter $\lambda_k$.
A convenient choice is to use the inverse Gamma distribution,
commonly parametrized in terms of the hyperparameters $a_k > 0$ and $b_k > 0$:
%Conjugate prior is $\lambda_k \sim \ig(a_k,b_k)$:
\begin{linenomath*}
\begin{align}
f_{\lambda_k}(\lambda_k\mid a_k,b_k) &= \frac{(b_k)^{a_k}}{\Gamma(a_k)} \lambda_k^{-a_k -1} e^{-\frac{b_k}{\lambda_k}}
\end{align}
\end{linenomath*}
in short, $\lambda_k \mid a_k, b_k \sim \ig(a_k,b_k)$.
The inverse Gamma is convenient because it is a conjugate prior,
i.e., the posterior obtained by Bayes' rule is again inverse Gamma and thus easily tractable;
the prior parameters only need to be updated to obtain the posterior parameters.

***drop $k$ for rest of this and next section?***

In the standard Bayesian approach, 
one has to fix a prior by choosing values for $a_k$ and $b_k$
to encode specific prior information about component lifetimes.
In our imprecise approach, we allow instead these parameters
to vary in a set, this is advantageous also
because expert knowledge is often vague,
and it is difficult for the expert(s) to pin down precise hyperparameter values.
For the definition of the hyperparameter set,
we use however a parametrization in terms of $\nz > 1$ and $\yz > 0$ instead of $a_k$ and $b_k$,
\begin{linenomath*}
\begin{align}
\nz &= \alpha - 1\,,
&
\yz &= \beta / \nz \,,
\end{align}
\end{linenomath*}
where $\yz$ can be interpreted as the prior guess for the scale parameter $\lambda$,
as $\E[\lambda\mid\nz,\yz] = \yz$.
This parametrization also makes the nature of the combination
of prior information and data through Bayes' rule more clear:
After observing $n$ component lifetimes $\vec{t} = (t_1, \ldots, t_n)$,
the updated parameters are
\begin{linenomath*}
\begin{align}
\nn &= \nz + n\,, 
&
\yn &=  \frac{\nz \yz + \taut}{\nz + n}\,,
\label{eq:ig-update}
\end{align}
\end{linenomath*}
where $\taut = \sum_{i=1}^n t_i^k$. %
%\footnote{
We thus have
\begin{linenomath*}
\begin{align}
\lambda \mid \nz, \yz, \vec{t} \sim \ig(\nz + n + 1, \nz \yz + \taut). %}
\label{eq:ig-update-alpha}
\end{align}
\end{linenomath*}
From the simple update rule \eqref{eq:ig-update}, we see that
$\yn$ is a weighted average of the prior parameter $\yz$ and the Maximum Likelihood estimator $\taut/n$,
with weights $\nz$ and $n$, respectively.
$\nz$ can thus be interpreted as a prior strength or pseudocount,
indicating how much our prior guess should weigh against the $n$ observations.
Furthermore, $\V[\lambda\mid\nz,\yz] = (\yz)^2 / (1 - 1/\nz)$,
so for fixed $\yz$, the higher $\nz$,
the more probability mass is concentrated around $\yz$. %$p(\lambda\mid\nz,\yz)$

%In terms of canonical parameters $\nkz$ and $\ykz$, we assume $\lambda_k \mid \nkz,\ykz \sim \ig(\nkz + 1, \nkz\ykz)$.
